Clustering and Parameter Estimation for the Popularity Adjusted Block Model

### 2020-07-14 summary.Rmd
minor updates to writeup

**Theorem 1**. 
Let $X = \begin{bmatrix} 
\lambda^{(11)} & \lambda^{(12)} & 0 & 0 \\ 
0 & 0 & \lambda^{(21)} & \lambda^{(22)} 
\end{bmatrix}$ 
where the $\lambda^{(kl)}$'s are defined as in Definition 1 for $K=2$, and 
let $U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
0 & 0 & 1 / \sqrt{2} & - 1 / \sqrt{2} \\
0 & 1 & 0 & 0 \end{bmatrix}$. 
Then $A \sim GRDPG_{3, 1}(3, 1)$ and $A \sim PABM(\{(\lambda^{(kl)}\}_K)$ are 
equivalent.

@@

*Proof*
Let $\Lambda^{(k)} = 
\begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
\in [0, 1]^{n_k \times K}$.  
Let $X$ be a block diagonal matrix 
$X = \diag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in [0, 1]^{n \times K^2}$.  
Let $L^{(k)}$ be a block diagonal matrix of column vectors $\lambda^{(lk)}$ for 
$l = 1, ..., K$. $L^{(k)} = \diag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
[0, 1]^{n \times K}$.  
Let $Y = \begin{bmatrix} L^{(1)} & \cdots & L^{(K)} \end{bmatrix} \in 
[0, 1]^{n \times K^2}$.  
Then $P = X Y^\top$.  
Similar to the $K = 2$ case, we again have $Y = X \Pi$ for a permutation matrix
$\Pi$, so $P = X \Pi X^\top$.  
The permutation described by $\Pi$ has $K$ fixed points, which correspond to 
$K$ eigenvalues equal to $1$ with corresponding eigenvectors $e_k$ where 
$k = r (K + 1) + 1$ for $r = 0, ..., K - 1$. It also has 
$\binom{K}{2} = K (K - 1) / 2$ cycles of order $2$. Each cycle corresponds to 
a pair of eigenvalues $+1$ and $-1$ and a pair of eigenvectors 
$(e_s + e_t) / \sqrt{2}$ and $(e_s - e_t) / \sqrt{2}$.

@@

In addition, $(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}$ converge to a generalized 
chi-square distribution.

>>>>

**Theorem 1**. 
Let $X = \begin{bmatrix} 
\lambda^{(11)} & \lambda^{(12)} & 0 & 0 \\ 
0 & 0 & \lambda^{(21)} & \lambda^{(22)} 
\end{bmatrix}$, and let $U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
0 & 0 & 1 / \sqrt{2} & - 1 / \sqrt{2} \\
0 & 1 & 0 & 0 \end{bmatrix}$, as in Definition 1. 
Then $A \sim GRDPG_{3, 1}(X U)$ and $A \sim PABM(\{(\lambda^{(kl)}\}_K)$ are 
equivalent.

@@

*Proof*
Let $\Lambda^{(k)} = 
\begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
\in \mathbb{R}^{n_k \times K}$.  
Let $X$ be a block diagonal matrix 
$X = \diag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}$.  
Let $L^{(k)}$ be a block diagonal matrix of column vectors $\lambda^{(lk)}$ for 
$l = 1, ..., K$. $L^{(k)} = \diag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
\mathbb{R}^{n \times K}$.  
Let $Y = \begin{bmatrix} L^{(1)} & \cdots & L^{(K)} \end{bmatrix} \in 
\mathbb{R}^{n \times K^2}$.  
Then $P = X Y^\top$.  
Similar to the $K = 2$ case, we again have $Y = X \Pi$ for a permutation matrix
$\Pi$, so $P = X \Pi X^\top$.  
The permutation described by $\Pi$ has $K$ fixed points, which correspond to 
$K$ eigenvalues equal to $1$ with corresponding eigenvectors $e_k$ where 
$k = r (K + 1) + 1$ for $r = 0, ..., K - 1$. It also has 
$\binom{K}{2} = K (K - 1) / 2$ cycles of order $2$. Each cycle corresponds to 
a pair of eigenvalues $+1$ and $-1$ and a pair of eigenvectors 
$(e_s + e_t) / \sqrt{2}$ and $(e_s - e_t) / \sqrt{2}$.

@@

In addition, by the central limit theorem, 
$(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}$ converge to a normal distribution 
centered at $0$.
